---
title: "Assignment 2 part 2"
output: pdf_document
---
1. What can you say about the central tendency of the ratings?
```{r}
zagat <- read.csv("https://raw.githubusercontent.com/jcbonilla/BusinessAnalytics/master/BAData/zagat.CSV")
summary(zagat)
apply(zagat[,2:5],2,mean) # apply function helps to implement one function to all variables
apply(zagat[,2:5],2,median)
apply(zagat[,2:5],2,function(x)names(sort(table(x),decreasing = T))[1])
```
For centeral tendency you can also generate box plots and histograms
```{r}
par(mfrow=c(2,2)) # 4 figures arranged in 2 rows and 2 columns
hist(zagat$Food,col="red")
hist(zagat$Decor, col="blue")
hist(zagat$Service,col="green")
hist(zagat$Price,col ="yellow")
```
Boxplots
```{r}
par(mfrow=c(4,1)) # 4 figues arranged in 4 rows and 1 column
boxplot(zagat$Food,horizontal = TRUE, col="red",xlab="Food")
boxplot(zagat$Decor,horizontal = TRUE, col="blue",xlab="Decor")
boxplot(zagat$Service,horizontal = TRUE, col="green",xlab="Service")
boxplot(zagat$Price,horizontal = TRUE, col="yellow",xlab="Price")
```

From the plots above we can see that Food is rightly skewed and food mean = median < mode
Decor is leftly skewed and mean = median > mode
Service is rightly skewed and median = mode < mean
Price is rightly skewed and mode< median< mean


2. What can you say about the spread and dispersion of the ratings?
```{r}
apply(zagat[,2:5],2,range) 
# The diff function computes the difference between pairs of consecutive elements of a numeric vector.
diff(apply(zagat[,2:5],2,range) )
#Quantiles
apply(zagat[,2:5],2,quantile)
#Interquartile
apply(zagat[,2:5],2,quantile)[4,]-apply(zagat[,2:5],2,quantile)[2,]
#Variance 
apply(zagat[,2:5],2,var)
#Standard Deviation
apply(zagat[,2:5],2,sd)
```

In all of these measures it can be observed clearly that price is a highly dispersed variable. Which indicates that customer ratings differ a lot when it comes to price. 

3. What are the correlations between rating dimensions?
```{r}
zagat_cor <- cor(zagat[,2:5])
zagat_cor

```
To get a visualization of correlation you can also generate heatmaps and correlation matrix
```{r}
#install.packages("tidyr")
library("tidyr")
#install.packages("ggplot2")
library("ggplot2")
#install.packages("gplots")
library("gplots")

rnames <- zagat[,1]                            # assign labels in column 1 to "rnames"
mat_data <- data.matrix(zagat[,2:ncol(zagat)])  # transform column 2-5 into a matrix
rownames(mat_data) <- rnames 
heatmap.2(mat_data, scale = "none", col = bluered(100), 
          trace = "none", density.info = "none")

library(corrplot)
par(mfrow=c(1,1))
corrplot(zagat_cor,type = "upper", order = "hclust", tl.col="black",tl.sort=90, addCoef.col = "white")


```

4. Using the information in 1-3, design a weighted average (index) that computes scores for each restaurant. Your index needs to reflect which ratings (decor vs food vs service vs price) you wish to amplify with loads/weights


```{r}
#standardizing the data
#Z-score formula

food_s <- (zagat$Food)/mean(zagat$Food)
decor_s <- (zagat$Decor)/mean(zagat$Decor)
service_s <- (zagat$Service)/mean(zagat$Service)
price_s <- (zagat$Price)/mean(zagat$price)

#sum of standard deviations
sd_sum = sd(food_s) + sd(decor_s) + sd(service_s) + sd(price_s)

#Calculating individual weights
food_weight= (range(food_s)[2]-range(food_s)[1])*sd(food_s)/sd_sum
decor_weight= (range(decor_s)[2]-range(decor_s)[1])*sd(decor_s)/sd_sum
service_weight= (range(service_s)[2]-range(service_s)[1])*sd(service_s)/sd_sum
price_weight= (range(price_s)[2]-range(price_s)[1])*sd(price_s)/sd_sum

#Standardizing weights
total_weight = food_weight+decor_weight+service_weight+price_weight

food_w = food_weight/total_weight
decor_w = decor_weight/total_weight
service_w = service_weight/total_weight
price_w = price_weight/total_weight

food_w 
decor_w 
service_w 
price_w 

```
Some Business Insights:

1. From the evaluation above it can be seen that price is highly dispersed and would impact any business model the most.
2. Price has least correaltion food hence I can say that improving the quality of food will help there business model. 